GeoSched: Minimizing Cost of Cloud Workloads in Geo-Distributed Data Centers

### 2015-04-05 doc/report/midterm/article.tex
tex + bib

%%\documentclass[]{spie}  %>>> use for US letter paper
\documentclass[12pt]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage[]{graphicx}
\usepackage[sort, numbers]{natbib}
\title{GeoSched: Minimizing Cost of Cloud Workloads in Geo-Distributed Data Centers} 
\author{Anirudh Jayakumar and Harshit Dokania \\
University of Illinois at Urbana-Champaign \\
\{ajayaku2,hdokani2\}@illinois.edu 
}

@@

@@

@@

\section{Cloud Workload }

\subsection{Workload Traces}

\subsection{Google Workload Trace}

\subsection{Custom Trace Format}

\section{Data Center Profiling }

@@

    \begin{tabular}{ | p{3cm} |l  |p{11cm}|  }
    \hline
    \bf{Item} & \bf{Deadline} & \bf{Description}   \\ \hline
    Study real-world and synthetic workloads & 10\textsuperscript{th} March,2015  & We have started studying workloads traces for google cluster-data which provides traces from a 12k-machine cell for one month duration in May 2011.  \\ \hline
    Data Center profile & 20\textsuperscript{th} March,2015  & In this process we will note the data center specific parameters such as electricity cost, available energy, economizer, cooling energy that is useful for simulating geo-distributed data centers. \\ \hline
    Simulator design & 5\textsuperscript{th} April,2015  & Implement the simulator infrastructure for multi-datacenter environment.  \\ \hline
    Feature Extraction & 8\textsuperscript{th} April,2015  & This phase will be identifying all the necessary features that contribute to the energy cost. These features will be used in the optimization of the energy cost \\ \hline
    Implementing Optimization technique & 25\textsuperscript{th} April,2015  & This is the most critical component and expected to consume major part of project. We will be building optimization techniques which is adaptive to workload characteristics and data center's resource capacity and electricity costs \\ \hline
    Evaluation & 7\textsuperscript{th} May,2015  & This phase we will be simulating data centers with previously gathered data center profile and using the optimization technique mentioned above \\ \hline

>>>>

%%\documentclass[]{spie}  %>>> use for US letter paper
\documentclass[12pt]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage[]{graphicx}
\usepackage[sort, numbers]{natbib}
\usepackage{hyperref}
\usepackage{adjustbox,lipsum}
\graphicspath{ {/Users/anirudhnair/code/GeoSched/src/GoogleTrace/plots} }
\usepackage{subfig}

\title{GeoSched: Minimizing Cost of Cloud Workloads in Geo-Distributed Data Centers} 
\author{Anirudh Jayakumar and Harshit Dokania \\
University of Illinois at Urbana-Champaign \\
\{ajayaku2,hdokani2\}@illinois.edu 
}

@@

\iffalse

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fi

\section{Cloud Workload } 
With the increase in the adoption of cloud computing by businesses around the world, more and more workloads are moving out of traditional data centers into the cloud. A cloud data center typically handles a range of IT workloads. These workloads include service-critical and latency sensitive interactive applications that run 24x7 such as instant messengers, e-mail applications and other internet services, and batch-style applications such as scientific HPC applications, BigData applications like financial analysis and MapReduce applications. A thoughtful management of these workloads are necessary not only to provide customers with low-latency services but also to reduce the energy consumption of the data centers that run these workloads. A good way to study these workloads are by analyzing real-world or synthetic workload traces. We look at workload traces in detail in section \ref{traces} and \ref{googletrace}

\subsection{Workload Traces} \label{traces}
Unfortunately, there is a lack of real-world traces in the cloud computing research community. As a result, there are limited work done to understand the diversity in cloud workloads. The first large-scale workload trace was released by Google in 2011\cite{googletracedata}, featuring traces from over 12,000 servers over a period of 29 days. Various studies have been carried out to understand the characteristic of the Google trace data. Other real-world traces include Wikipedia request trace \cite{urdaneta2009wikipedia} and Hadoop logs from Yahoo! \cite{yahootrace}There have also been attempts to generate synthetic workload traces using known real-world workload characteristics \cite{beitch2010rain} \cite{wang2011towards}. For the sake of this work, we have decided to use the workload trace from Google.

\subsection{Custom Trace Format}
To have a common trace format to represent different real-world and synthetic workload traces, we have formulated a trace format similar to standard workload format (swf) \cite{chapin1999benchmarks} defined for grid and supercomputing traces. Table \ref{tab:format} provides information about each field int the trace format.

\begin {table}[ht]
\caption {Trace format description} \label{tab:format} 
 \scalebox{0.9}{\begin{tabular}{ | p{4cm} |l  |p{11 cm}|  }
    \hline
    \bf{Field Name} & \bf{Type} & \bf{Description}   \\ \hline
     JobID & INTEGER  & Unique job identifier for each job\\ \hline
     Arrival Time & INTEGER & Arrival time in microseconds \\ \hline
     Scheduled Time & INTEGER  & Arrival time in microseconds\\ \hline
     Finish Time & INTEGER  & Arrival time in microseconds\\ \hline
     Schedule class & INTEGER  & Latency class of the job. \\ \hline
     Tasks & INTEGER  & Total number of tasks in the job. \\ \hline
     Total CPU& INTEGER  & Aggregate sum of CPU used by all the tasks in the job\\ \hline
     Total Memory& INTEGER   & Aggregate sum of memory used by all the tasks in the job \\ \hline
     User& INTEGER  & Unique user identifier \\ \hline
     SLA Param1&  FLOAT & SLA parameter placeholder\\ \hline
     SLA Param2&  FLOAT & SLA parameter placeholder\\ \hline
     SLA Param3&  FLOAT & SLA parameter placeholder\\ \hline
    \end{tabular}}
\end{table}

\subsection{Google Workload Trace} \label{googletrace}
Google trace has job and task information in different files. We consolidated all the required information and generated a single work trace file with the format described in the previous section. For our work, we model five different data centers located at different parts of the globe. A detailed explanation about the data center profile are given in section \ref{sec:datacenter}. To simulate workload execution in these five data centers we have extracted five set of traces each worth five days of incoming workload from the original google trace. So, this covers 25 days of traces out of the 29 available days. We label these 5 set of traces from A to E. Additionally, in our experiments we wish to only consider those jobs that end without any failures. Therefore, we clean the original traces by removing all jobs that is either failed or get killed.

\begin{figure}[ht] 
\centering
\includegraphics[scale=0.40]{arrivalrate.png}
\caption{Arrival rate of jobs }
\label{fig:arrivalrate}
\end{figure}

Now we will look at some characteristics of the traces. Table \ref{tab:tracestat} describes the statistics for each of the five traces grouped by the scheduling class of the job. In google traces, CPU and memory usage are normalized relative to the largest capacity of the resource on any machine in the trace (which is 1.0). Most of the class 0 and class 2 jobs are batch jobs with less latency requirements. Generally, these are the kind of jobs that can be provisioned on other data centers as their deadlines are not strict.  You will find that as the sensitivity increases the running time of these jobs also increase. High-sensitive jobs are generally service jobs and run of longer periods of time. Figure \ref{fig:pie} shows a pie chart of the job counts in different schedule classes for all the 5 traces. You will notice that the job count for class 3 (high latency-sensitivity) jobs are low. This is because these jobs generally run for long period of time, typically months. Since we only consider jobs that start and end within the 29 day range such jobs are missed out in our traces. For our work, we will consider class-0 and class-1 jobs as candidates for optimizing the provisioning. We leave, class-2 and class-3 jobs to run on the data center where the job arrived since these are high-sensitive and their SLA demands will be strict.

\begin{figure}[]
\centering
\includegraphics[scale=0.65]{pie}
\caption{ Distribution of job count in various schedule classes}
 \label{fig:pie}
\end{figure}

\begin{table}
\caption {Trace set statistics} \label{tab:tracestat} 
\adjustbox{max width=1\textwidth}{

\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|l|}
\hline
 & \multicolumn{3}{l|}{\bf{Trace A}} & \multicolumn{3}{l|}{\bf{Trace B}} & \multicolumn{3}{l|}{\bf{Trace C}} & \multicolumn{3}{l|}{\bf{Trace D}} & \multicolumn{3}{l|}{\bf{Trace E}} \\ \hline
\textbf{\begin{tabular}[c]{@{}l@{}}Sched\\ Class\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg \\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ CPU\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}Avg\\ Mem\end{tabular}} & \textbf{\begin{tabular}[c]{@{}l@{}}RT\\  (secs)\end{tabular}} \\ \hline
0 & 0.075 & 0.051 & 415 & 0.123 & 0.096 & 345 & 0.083 & 0.065 & 385 & 0.093 & 0.092 & 373 & 0.093 & 0.074 & 333 \\ \hline
1 & 0.087 & 0.029 & 729 & 0.086 & 0.032 & 760 & 0.082 & 0.028 & 711 & 0.073 & 0.030 & 743 & 0.093 & 0.043 & 693 \\ \hline
2 & 0.067 & 0.043 & 1028 & 0.096 & 0.058 & 872 & 0.088 & 0.057 & 918 & 0.073 & 0.045 & 707 & 0.084 & 0.049 & 832 \\ \hline
3 & 0.088 & 0.12 & 1105 & 0.099 & 0.12 & 82 & 0.101 & 0.012 & 7961 & 0.081 & 0.051 & 17135 & 0.089 & 0.061 & 9021 \\ \hline
\end{tabular}
}
\end{table}

Figure \ref{fig:arrivalrate} shows the variation in job arrival for the traces. As you can see there is both intra-trace variation as well as inter-trace variation. This is representative of the real-world cloud workload pattern where bursty arrivals are common. There are opportunities to provision jobs to other data centers if the traffic at a particular data center is high. What is also important to notice is that the scheduling decision should be made quickly when the arrival rates of jobs are high, else there are chances for these jobs to miss the SLA. Hence, even the scheduler should be scalable during high volume traffic.

Figure \ref{fig:runtime} shows the CDF of the job runtime. As you can see most of the jobs have runtime less than 1024 seconds. This pattern is consistent for all the trace sets. There are a few jobs that have higher run-times. As explained earlier, these jobs are primarily internet-service jobs that run for months.

\begin{figure}[] 
\centering
\includegraphics[scale=0.40]{runtime}
\caption{CDF of the job duration}
\label{fig:runtime}
\end{figure}

In google workload trace, it is tough to study the resource usage since these value are scaled down and therefore and accurate picture of the resource utilization is tough to generate. In figure \ref{fig:resource} we can see that most jobs request a small portion of the cluster resources. Most of the resources in the google cluster are used by the few service jobs that are present. The large number of batch jobs only use a small percentage of the resources \cite{mishra2010towards} \cite{reiss2012heterogeneity}. From the above statistics, we believe that google workload trace provides an accurate representation of the real-world cloud datacenter workload.

\begin{figure} %
    \centering
    \subfloat[CPU resource ]{{\includegraphics[width=9cm]{cpu} }}%
    \subfloat[RAM resource]{{\includegraphics[width=9cm]{mem} }}%
    \caption{CDF of resource requested by jobs}%
    \label{fig:resource}%
\end{figure}

@@

\section{Data Center Profiling } \label{sec:datacenter}

@@

    \begin{tabular}{ | p{4cm} |l  |p{11 cm}|  }
    \hline
    \bf{Item} & \bf{Deadline} & \bf{Description}   \\ \hline
    Study real-world and synthetic workloads & 10\textsuperscript{th} March  & We have started studying workloads traces for google cluster-data which provides traces from a 12k-machine cell for one month duration in May 2011.  \\ \hline
    Data Center profile & 20\textsuperscript{th} March  & In this process we will note the data center specific parameters such as electricity cost, available energy, economizer, cooling energy that is useful for simulating geo-distributed data centers. \\ \hline
    Simulator design and implementation& 12\textsuperscript{th} April  & Implement the simulator infrastructure for multi-datacenter environment.  \\ \hline
    Feature Extraction and Optimization & 25\textsuperscript{th} April  & Here we identifying all the necessary features that contribute to the energy cost  and build an optimization technique which is adaptive to workload characteristics and data center's resource capacity and electricity costs \\ \hline
    Evaluation & 7\textsuperscript{th} May  & This phase we will be simulating data centers with previously gathered data center profile and using the optimization technique mentioned above \\ \hline
